### 2021-04-21 paper_analysis.Rmd
actualizacion
labels prep data
tablas para paper

---
title: "Paper analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Descriptivos

## Univariado 
```{r echo=FALSE, results='asis',message=F, error=F}
pacman::p_load(summarytools,haven,sjPlot,dplyr,htmlwidgets)
load(file = "input/data/ap_est.RData")
basereg <- ap_est %>% select(perc_trabajo_duro_factor_est,
                             perc_esfuerzo_factor_est,
                             perc_trabajo_duro_factor_ap,
                             perc_esfuerzo_factor_ap,
                             sj_direct,
                             quintiles_ingresos_pc_factor,
                             libros_hogar_factor_ap,
                             recompensa_factor_est,
                             resp_prof_factor,
                             pos_pol_factor_ap,
                             pos_pol_factor_est,
                             genero_factor_est,
                             dependencia_factor_ap,
                             # rbd_est
                             ) %>% na.omit(basereg)
view(dfSummary(basereg,
               plain.ascii = FALSE,
               style = "grid",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = T, # plot
               valid.col = T, # n valido
               # col.widths = c(5,5,10,10,10,10)
               ),
     method = "render")
```

## Bivariado

```{r echo=FALSE}
pacman::p_load(summarytools,haven,sjPlot,dplyr,sjlabelled,rvest,knitr,kableExtra)
load(file = "input/data/ap_est.RData")
basereg <- ap_est %>% select(perc_trabajo_duro_factor_est,
                             perc_esfuerzo_factor_est,
                             perc_trabajo_duro_factor_ap,
                             perc_esfuerzo_factor_ap,
                             sj_direct,
                             quintiles_ingresos_pc_factor,
                             libros_hogar_factor_ap,
                             recompensa_est,
                             resp_prof_factor,
                             pos_pol_factor_ap,
                             pos_pol_factor_est,
                             genero_factor_est,
                             dependencia_factor_ap,
                             # rbd_est
                             ) %>% na.omit(basereg)

cor.df<- data.frame(lapply(basereg[,c("perc_trabajo_duro_factor_est",
                                       "perc_esfuerzo_factor_est",
                                       "perc_trabajo_duro_factor_ap",
                                       "perc_esfuerzo_factor_ap",
                                       "sj_direct")], function(x) as_numeric(x))) 
cor.label <- as.character(get_label(cor.df))
cor.tab<-sjt.corr(cor.df,
                  triangle = "lower",
                  digits = 2,
                  string.diag = rep(x = "$-$",max(length(cor.df))))
cor.mat <- as.data.frame(read_html(x = cor.tab[["page.content"]]) %>% 
                           html_table(fill=FALSE))
cor.mat <- cor.mat[-max(length(cor.mat)),] 
cor.mat$Var.1 <- paste0(1:length(cor.label),". ",cor.label)
colnames(cor.mat) <- as.character(c(" ",1:length(cor.label)))

kable(cor.mat,
      align = c("l",rep(x = "c",max(length(cor.df)))),
      caption = "Correlaciones de Pearson" ) %>% 
  kable_classic(full_width = T) %>%  
  footnote(paste0("$N=$ ",dim(basereg)[1],", ","$^{***}$p< 0.001, $^{**}$p< 0.01, $^{*}$p< 0.05"),
           footnote_as_chunk = T,
           general_title = "Nota: " )
```

# Modelos