### 2023-06-07 presentacion-effort/presentation-effort.Rmd

---
title: "The Socialization of Meritocracy: The role of the Family and the School"
subtitle: ".medium[**Resultados preliminares**]"
author: "Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas"
institute: ".small[**FONDECYT N°1210847 Meritocracia en la escuela**]"
output:
  xaringan::moon_reader:
    css: ["custom_lisa.css"]
    seal: false
    nature:
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "xaringan_custom/macros.js"
      ratio: '16:9'
---

class: middle title-slide

```{r xaringanExtra, echo = FALSE}
  xaringanExtra::use_progress_bar(color = "red", location = c("top"))
```

.pull-left-narrow[

<br>
<br>
<br>
<br>
<br>
<br>
<br>

![:scale 70%](images/paces.png)
.small[.red[FONDECYT N°1181239 Political socialisation and citizenship education]
]
]

.pull-right-wide[
.right[

.content-box-red[
## The Socialization of Meritocracy: The role of the Family and the School

]

----
<br>
.espaciosimplelineas[
.medium[
Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas
] 
]

.small[Effort and Inequality Conference - 16th June 2022]
]
]

---

layout: true
class: animated, fadeIn

---
class: middle

.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar
]

.pull-right-wide[
![](images/price.png)

.right[
[ciudadania-escolar.cl/](https://ciudadania-escolar.cl)
]]

---
class: middle

.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar
]

.pull-right-wide[

![](images/proyectos.png)

.right[
Projects since 2011 ]
]

---
.pull-left-narrow[
# **PRICE**:
### Programa de Investigación en Ciudadanía Escolar

(School Citizenship Research Programme)
]

.pull-right-wide[

![](images/equipo.png)
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. Results
5. Discussion
]
]

---
class: inverse

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. .yellow[Introduction]
2. Hypothesis
3. Methodology
4. Results
5. Discussion
]
]

---

# .black[Introduction]

.pull-left[
  .content-box-gray[
## Meritocracy
  - Effort and talent  
  - Descriptive (perception) 
  - Normative (preferences)   
  ] 
]

--
.pull-right[
<br><br>

.content-box-green[
## Socialisation
  - Family and school as socialisation agents 
  - Justice experiences at school
  ]  
]

---
class:inverse middle left

###.yellow[
_Is there an association between the meritocratic perceptions of parents with those of their offsprings?_
]

### .yellow[_To what extent is school experience in terms of distributive justice (in relation to grades) linked to the meritocratic perceptions of students?_]

---
# .black[Introduction]

----

.pull-left[
.content-box-purple[
## Family

  - Socio-economic status (self-interest)

  - Intergenerational transmission of beleifs and attitudes (socialisation)

<br>

  ]
  ]

.pull-right[
.content-box-yellow[
## School

- Meritocracy perceptions (school and society)

- Justice evaluations about grades

<br>

]

]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]
.pull-right[
.large[
1. Introduction
2. .yellow[Hypothesis]
3. Methodology
4. Results
5. Discussion
]
]

---
## Hypothesis

.center[![:scale 70%](images/model-hip-eng.PNG)]

.small[
* Pre-registration: [https://osf.io/fazdx](https://osf.io/fazdx)
]
---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. .yellow[Method]
4. Results
5. Discussion
]
]

---
## Data and methods
----

.pull-left[

* School Citizenship Panel Survey (2019)

* Students, parents and teachers

* Analytical sample ( $n_i$ =541)

  - Student and parent information
  - $n_j$ = 60 (schools)

* _Method:_  Multilevel Ordinal logit  
]

.pull-right[
.right[![:scale 200%](images/paces.png)
]
]

---

### DV: Perceived Meritocracy

.center[![:scale 78%](images/perc-merit.jpg)]

---
### IV: Justice in grades

.pull-left[

*Indirect*

$$\text{Justice}= \ln\biggl(\frac{\text{grade obtained}}{\text{grade deserved}}\biggl)$$

.center[![:scale 93%](images/indirect-justice.png)]

]

.pull-right[

*Direct*

"Taking into account the time I devote to my studies, the grades I get are:"

.center[![:scale 100%](images/direct.png)]

  ]

---

## Dependent variables
----
.pull-left[
.content-box-purple[
### Family

* Educational level of parents

* Number of books in the household

* Household Income Quintile

]
]

.pull-right[
.content-box-yellow[
### School

* Type of School (Public, Charter, Private)

* School Socioeconomic Level (Adm. data)

]
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. .yellow[Results]
5. Discussion
]
]
---

.pull-left-narrow[
<br>
### How important is hard work to get ahead in life?

]

.pull-right-wide[
.center[![:scale 85%](images/plotgetahead.png)]
]
---

.pull-left-narrow[
<br>
### How important is hard work to get ahead in life?
]

.pull-right-wide[
.center[![:scale 85%](images/intergetahead.png)]
]

---

.pull-left-narrow[
<br>
### In Chile those who work hard get ahead
]

.pull-right-wide[

.center[![:scale 85%](images/ploteffort.png)]
]

---

.pull-left-narrow[
<br>
### In Chile those who work hard get ahead
]

.pull-right-wide[
.center[![:scale 85%](images/intereffort.png)]
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. Results
5. .yellow[Discussion]
]
]

---
# Discussion

- Family socialisation: mixed, relevant only for the importance of hard work
- SES-family has an inconsistent association. SES at the school level positively influences the importance of hard work but is null for perceived rewards.

- Justice experience in grades .bold[does not] have a direct influence on perceived meritocracy

- Interactions are mixed, decreasing the influence of socialisation on the importance of hard work, but positive effort is rewarded.

---
class: roja middle

# ¡Thank you!

.right[
.yellow[More information:

[www.ciudadania-escolar.cl](https://ciudadania-escolar.cl/)
]]

---
class: middle title-slide

.pull-left-narrow[

<br>
<br>
<br>
<br>
<br>
<br>
<br>

![:scale 70%](images/paces.png)
.small[.red[FONDECYT N°1181239 Political socialisation and citizenship education]
]
]

.pull-right-wide[
.right[

.content-box-red[
## The Socialization of Meritocracy: The role of the Family and the School

]

----
<br>
.espaciosimplelineas[
.medium[
Juan Carlos Castillo, Julio Iturra, Francisco Meneses & Martín Venegas
] 
]

.small[Effort and Inequality Conference - 16th June 2022]
]
]