### 2023-06-16 docs/presentacion-effort/presentation-effort.Rmd
ajustes slides

---
title: "The Socialization of Meritocracy: The role of the Family and the School"
subtitle: ".medium[**Resultados preliminares**]"
author: "Julio Iturra, Juan Carlos Castillo, Francisco Meneses & Martín Venegas"
institute: ".small[**FONDECYT N°1210847 Meritocracia en la escuela**]"
output:
  xaringan::moon_reader:
    css: ["custom_lisa.css"]
    seal: false
    nature:
      highlightStyle: solarized-dark
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "xaringan_custom/macros.js"
      ratio: '16:9'
---

class: middle title-slide

```{r xaringanExtra, echo = FALSE}
  xaringanExtra::use_progress_bar(color = "red", location = c("top"))
```

.pull-left-narrow[

<br>
<br>
<br>
<br>
<br>
<br>
<br>

![:scale 70%](images/paces.png)
.small[.red[FONDECYT N°1181239 Political socialisation and citizenship education]
]
]

.pull-right-wide[
.right[

.content-box-red[
## The Socialization of Meritocracy: The role of the Family and the School

]

----
<br>
.espaciosimplelineas[
.small[
Julio Iturra -  Bremen Int. Graduate School of Social Sciences
[jiturra@bigsss.uni-bremen.de](jiturra@bigsss.uni-bremen.de)
<br>
Juan Carlos Castillo -  Universidad de Chile
<br>
Francisco Meneses & Martín Venegas
] 
]

.small[Effort and Inequality Conference - 16th June 2022]
]
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. Results
5. Discussion
]
]

---
class: inverse

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. .yellow[Introduction]
2. Hypothesis
3. Methodology
4. Results
5. Discussion
]
]

---
<br>
.pull-left[
  .content-box-gray[
## Meritocracy
  - Merit: .bold[Effort] and talent (Young, 1958)  
  - Subjective meritocracy (Duru-Bellat & Tenret, 2012;Castillo et al, 2019)
    - Normative (_How it should be_) 
    - **Perception** (_How is it_) 
  ] 
]

.pull-right[
  .content-box-purple[
## Justice experiences
  - Justice experiences in school (Jasso & Resh, 2002)
  - Deserved vs obtained .bold[rewards]  (Resh & Sabbagh, 2014)
  ] 
]

---
# .black[Socialization agents]
----

.pull-left[
.content-box-purple[
## Family

  - Self-interest: family socio-economic resources

  - Intergenerational transmission of beliefs and attitudes (Bandura et al., 2001; Olivos, 2021)

  ]
  ]

.pull-right[
.content-box-yellow[

## School

- Meritocracy perceptions (Lampert, 2013; Mijs, 2018).

- Direct and indirect evaluation about grades (Resh, 2018; Resh & Sabbagh, 2014)

<br>

]

]

---
class:inverse middle left

###.yellow[.small[_What is the association between parental perceived meritocracy with their children's perceived meritocracy?_]]

### .yellow[.small[_To what extent the experiences of justice in grades influence student's perceived meritocracy?_]]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]
.pull-right[
.large[
1. Introduction
2. .yellow[Hypothesis]
3. Methodology
4. Results
5. Discussion
]
]

---
## Hypothesis

.center[![:scale 70%](images/model-hip-eng.PNG)]

.small[
* Pre-registration: [https://osf.io/fazdx](https://osf.io/fazdx)
]
---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. .yellow[Method]
4. Results
5. Discussion
]
]

---
## Data and methods
----

.pull-left[

* School Citizenship Panel Survey (2019)

- High school students (2nd) and parents (n = 541) nested in schools (n = 60)

* _Method:_  Multilevel regression models   
]

.pull-right[
.right[![:scale 200%](images/paces.png)
]
]

---

### Perceived Meritocracy

_Today in Chile, for getting ahead in life, how important is (...) Hard work_

<br>

.center[
![:scale 100%](images/perc-merit.jpg)
]

---
###  Justice in grades: Indirect

.pull-left[
.bold[Obtained:] _What was your last year average grade? If you don't remember exactly, please give an approximate grade._ (M= 5.8)

.bold[Deserved:] _And, what grade average do you think you deserved?_ (M=6.0)

Min: 1 / Max: 7

$$\text{Justice}= \ln\biggl(\frac{\text{grade obtained}}{\text{grade deserved}}\biggl)$$

]

.pull-right[

.center[![:scale 100%](images/indirect-justice.png)]

]

---
### Justice in grades:  Direct

_Taking into account the time I dedicate to my studies, the grades I get are:_

.center[![:scale 90%](images/direct.png)]

---

## Dependent variables
----
.pull-left[
.content-box-purple[
### Family

* Educational level of parents

* Number of books in the household

* Household Income Quintile

]
]

.pull-right[
.content-box-yellow[
### School

* Type of School (Public, Charter, Private)

* School Socioeconomic Level (Adm. data)

]
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. .yellow[Results]
5. Discussion
]
]
---

.pull-left-narrow[
<br>
### How important is hard work to get ahead in life?

]

.pull-right-wide[
.center[![:scale 85%](images/plotgetahead.png)]
]
---

.pull-left-narrow[
<br>
### How important is hard work to get ahead in life?
]

.pull-right-wide[

.center[![:scale 85%](images/intergetahead1.png)]
]

---
class: inverse animated slideInRight

## .red[The Socialization of Meritocracy: The role of the Family and the School
]

.pull-right[
.large[
1. Introduction
2. Hypothesis
3. Methodology
4. Results
5. .yellow[Discussion]
]
]

---
# Discussion

- Family socialisation: parents positively influence student's perceived meritocracy

- SES-family has an inconsistent association. SES at the school does not influence student's perceived meritocracy in terms of the importance of hard work
- Justice experience in grades .bold[does not] have a direct influence on perceived meritocracy

- Negative justice experiences in grades descrease the influence of socialisation on the importance of hard work.

---
class: roja middle center

# Thank you!

.right[
.yellow[More information:

[jiturra@bigsss.uni-bremen.de](jiturra@bigsss.uni-bremen.de)

[github.com/juancarloscastillo/socialisacion-merit](https://github.com/juancarloscastillo/socialisacion-merit)

[www.ciudadania-escolar.cl](https://ciudadania-escolar.cl/)
]]

---
class: middle title-slide

.pull-left-narrow[

<br>
<br>
<br>
<br>
<br>
<br>
<br>

![:scale 70%](images/paces.png)
.small[.red[FONDECYT N°1181239 Political socialisation and citizenship education]
]
]

.pull-right-wide[
.right[

.content-box-red[
## The Socialization of Meritocracy: The role of the Family and the School

]

----
<br>
.espaciosimplelineas[
.medium[
Julio Iturra -  Bremen Int. Graduate School of Social Sciences
[jiturra@bigsss.uni-bremen.de](jiturra@bigsss.uni-bremen.de)
<br>
Juan Carlos Castillo -  Universidad de Chile
<br>
Francisco Meneses & Martín Venegas
] 
]

.small[Effort and Inequality Conference - 16th June 2022]
]
]

---

![:scale 35%](images/descriptivos1.png)